---
title: "p8105_hw1_yc3242"
author: "Youn Kyeong Chang (uni# yc3242)"
date: "2018-09-17"
output: github_document
---

We use the tidyverse library through the whole problem set.
```{r library}
library(tidyverse)
```

## Problem 1

### Create data frame 
```{r create_df_p1}
df_p1 = tibble(
  vec_num = runif(10, min = 0, max = 5),
  vec_logi = vec_num > 2,
  vec_char = c("j", "O", "O", "N", "h", "e", "e", "K", "W", "n"),
  vec_fac = as.factor(c("upper", "lower", "lower", "upper", "upper", "lower", 
                     "upper", "upper", "lower", "upper"))
)
```

### Take the mean of each variable 
```{r mean_p1, warning = FALSE}
mean(df_p1$vec_num)
mean(df_p1$vec_logi)
mean(df_p1$vec_char)
mean(df_p1$vec_fac)
```
The mean of numeric and logical vectors can be calculated since they are numerical variables. 
However, that of character and factor vectors cannot be taken because they are categorical variables.

### Convert variable types to numeric at once
```{r convert_once_p1, eval = FALSE}
as.numeric(df_p1$vec_logi)
as.numeric(df_p1$vec_char)
as.numeric(df_p1$vec_fac)
```
Logical and factor vectors were able to be converted to numeric vectors but character vector was not.

### Convert variable types to numeric using steps
```{r convert_twice_p1, warning = FALSE}
as.numeric(as.factor(df_p1$vec_char))
as.numeric(as.character(df_p1$vec_fac))
```
Character vector was converted to numeric vector. However, factor vector was not converted to numeric vector.


## Problem 2

### Create data frame
```{r create_df_p2}
df_p2 = tibble(
  x = rnorm(1000),
  y = rnorm(1000),
  pos_xy = x + y > 0,
  num_pos_xy = as.numeric(pos_xy),
  fac_pos_xy = as.factor(pos_xy)
)
```
* The **size** of the dataset df_p2 is `r dim(df_p2)`, which means the dataset has 5 vectors of the length 1000. 
* The **mean** and the **median** of **x** is `r round(mean(df_p2$x), 2)` and `r round(median(df_p2$x), 2)`, respectively. 
* The **mean** and the **median** of **y** is `r round(mean(df_p2$y), 2)` and `r round(median(df_p2$y), 2)`, respectively. 
* The **proportion** of cases for which the logical vector is _TRUE_ is `r round(sum(df_p2$num_pos_xy)/nrow(df_p2), 2)`. 

### Visualize data with scatterplots of y vs x
```{r library_p2}
library(ggplot2)
```

```{r scatterplot_p2}
pos_xy_scatter = ggplot(df_p2, aes(x = x, y = y, color = pos_xy)) + geom_point()
num_pos_xy_scatter = ggplot(df_p2, aes(x = x, y = y, color = num_pos_xy)) + geom_point()
fac_pos_xy_scatter = ggplot(df_p2, aes(x = x, y = y, color = fac_pos_xy)) + geom_point()
```
To evaluate the binary value for ``x + y > 0``, while the logical vector ``pos_xy`` and the factor vector `` fac_pos_xy`` have the same binary color scale, the numeric vector ``num_pos_xy`` has different binary colors with a gradient legend.

### Save a scattorplot
```{r}
ggsave("scatterplot_p2.pdf", plot = pos_xy_scatter)
```

