---
title: "p8105_hw1_yc3242"
author: "Youn Kyeong Chang (uni# yc3242)"
date: "2018-09-17"
output: github_document
---

We use the tidyverse library through the whole problem set.
```{r library}
library(tidyverse)
```

# Problem 1

```{r p1_create_df}
df_p1 = tibble(
  vec_num = runif(10, min = 0, max = 5),
  vec_logi = vec_num > 2,
  vec_char = c("j", "O", "O", "N", "h", "e", "e", "K", "W", "n"),
  vec_fac = as.factor(c("upper", "lower", "lower", "upper", "upper", "lower", 
                     "upper", "upper", "lower", "upper"))
)
```

```{r p1_mean}
mean(df_p1$vec_num)
mean(df_p1$vec_logi)
mean(df_p1$vec_char)
mean(df_p1$vec_fac)
```
The mean of numeric and logical vectors can be calculated in that they are numerical variables. 
However, that of character and factor vectors cannot be taken because they are categorical variables.

```{r p1_convert, eval = FALSE}
as.numeric(df_p1$vec_logi)
as.numeric(df_p1$vec_char)
as.numeric(df_p1$vec_fac)
```
Logical and factor vectors were converted to numeric vectors but character vector failed.

```{r p1_char_convert}
as.numeric(as.factor(df_p1$vec_char))
```
Character vector was converted to numeric vector. 

# Problem 2

```{r p2_create_df}
df_p2 = tibble(
  x = rnorm(1000),
  y = rnorm(1000),
  pos_xy = x + y > 0,
  num_pos_xy = as.numeric(pos_xy),
  fac_pos_xy = as.factor(pos_xy)
)
```

The **size** of the dataset df_p2 is `r dim(df_p2)`. In this dataset, the **mean** and the **median** of **x** is `r round(mean(df_p2$x), 2)` and `r round(median(df_p2$x), 2)`, respectively. Also, the **mean** and the **median** of **y** is `r round(mean(df_p2$y), 2)` and `r round(median(df_p2$y), 2)`, respectively. The **proportion** of cases for which the logical vector is _TRUE_ is `r round(sum(df_p2$num_pos_xy)/nrow(df_p2), 2)`. 

